---
title: "Giulio"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)
```



```{r, include = FALSE, message= FALSE}
library(tidymodels)
library(tidyverse)
library(viridis)
library(stacks)
library(maps)
library(vip)
path <- 'dati_vino_regioni.csv'
vini <- read.csv(path)

```


## Explore data

Here we do a little pre-processing of our data and we do a PCA in order to see 5 main profiles of wines.

```{r data exploration}
mean_tilde <- function(word){
  if(is.na(word)) return(NA)
  if(!str_detect(word, '~')) return(as.numeric(word))
  vec <- as.numeric(unlist(str_split(word, '~')))
  return(mean(vec))
}

closest_uv <- function(tib, row){
  id_ <- row$id
  tib <- tib %>%
    filter(!is.na(uvindex)) %>%
    filter(id != id_) %>%
    arrange(abs(latitudine-row$latitudine)+abs(longitudine-row$longitudine)) %>%
    slice_head(n = 3)
  val <- weighted.mean(tib$uvindex, c(0.95, 0.04, 0.01))
  return(val)
}
closest_energy <- function(tib, row){
  id_ <- row$id
  tib <- tib %>%
    filter(!is.na(solarenergy)) %>%
    filter(id != id_) %>%
    arrange(abs(latitudine-row$latitudine)+abs(longitudine-row$longitudine)) %>%
    slice_head(n = 3)
  val <- weighted.mean(tib$solarenergy, c(0.95, 0.04, 0.01))
  return(val)
}
closest_radiation <- function(tib, row){
  id_ <- row$id
  tib <- tib %>%
    filter(!is.na(solarradiation)) %>%
    filter(id != id_) %>%
    arrange(abs(latitudine-row$latitudine)+abs(longitudine-row$longitudine)) %>%
    slice_head(n = 3)
  val <- weighted.mean(tib$solarradiation, c(0.95, 0.04, 0.01))
  return(val)
}


vini <- vini %>%
  mutate(
    price = ifelse(price == 0, NA, price),
    year = ifelse(year == 0, NA, year),
    uvindex = ifelse(uvindex == 0, NA, uvindex),
    solarenergy = ifelse(solarenergy == 0, NA, solarenergy),
    solarradiation = ifelse(solarradiation == 0, NA, solarradiation),
    abv = Vectorize(mean_tilde)(abv),
    degree = Vectorize(mean_tilde)(degree),
    degree = ifelse(degree==0, median(degree), degree),
    year = as.factor(year),
    local1 = as.factor(local1)
  ) %>%
  filter(
    tempmax > 0
  )

for(idx in seq_along(vini$id)) {
  if(is.na(vini$uvindex[idx])){
    vini$uvindex[idx] = closest_uv(vini, vini[idx,])
  }
  if(is.na(vini$solarenergy[idx])){
    vini$solarenergy[idx] = closest_energy(vini, vini[idx,])
  }
  if(is.na(vini$solarradiation[idx])){
    vini$solarradiation[idx] = closest_radiation(vini, vini[idx,])
  }
}


pca_rec <- na.omit(vini) %>%
  recipe(score~., data = .) %>%
  update_role(name, id = "id") %>%
  step_normalize(all_numeric_predictors()) %>%
  step_pca(abv:tannin, cec:uvindex, abv:tannin)

pca_prep <- prep(pca_rec)

tidied_pca <- tidy(pca_prep, 2)



```



```{r, echo = F}
tidied_pca %>%
  filter(component %in% paste0("PC", 1:5)) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(value, terms, fill = terms)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~component, nrow = 1) +
  labs(y = NULL)
```

```{r, echo = F} 
juice(pca_prep) %>%
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color = score), alpha = 0.7, size = 2) +
  labs(color = NULL)  + 
  scale_color_viridis(option="magma")+
  scale_fill_viridis() 
```


The principal components don't offer much ordered view of our data and the score is pretty uniformly distributed. 

Let's see if the spacial coordinated offer us a clearer view of what is happening.


```{r, echo = F}

Italy <- map_data('italy', region= '.', exact= F)

ggplot(Italy, aes(long, lat))+
  geom_polygon(aes(group = group, ), fill = '#4E84C4') +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)+
  geom_point(data = vini, aes(x=longitudine, y=latitudine, color = score), size=1)+
  scale_color_viridis(option="magma") +
  scale_fill_viridis() 
```


Let's observe, in detail, how the most important features are distributed. Below, we have violin plots of chemical components found in the soil and of weather conditions

```{r, echo = F}

vini %>%
  pivot_longer(cols=cec:silt,values_to = "obs", names_to = "grp") %>%
  ggplot(aes(x=grp,y=obs, fill=grp))+
  geom_violin()+
  ggforce::geom_sina(alpha=0.5, size=.015)+
  theme(legend.position="none")+
  ggtitle('Violin plot of the chemical composition of the soil') +
  xlab('Component') +
  ylab('Observations')
```

```{r, echo = F}

vini %>%
  pivot_longer(cols=c(tempmax:humidity, windspeed, solarenergy, uvindex),values_to = "obs", names_to = "grp") %>%
  ggplot(aes(x=grp,y=obs, fill=grp))+
  geom_violin()+
  ggforce::geom_sina(alpha=0.5, size=.015)+
  theme(legend.position="none") +
  ggtitle('Violin plot of the weather') +
  xlab('Weather Variables') +
  ylab('Observations')
```


## Build models

```{r}
set.seed(123)
vini_split <- 
  vini %>%
  select(latitudine, longitudine, cec:uvindex, abv:tannin, year, price, score) %>%
  na.omit() %>%
  initial_split(., prop = .75, strata = score)
vini_train <- training(vini_split)
vini_test <- testing(vini_split)

set.seed(234)
vini_folds <- vfold_cv(vini_train, strata = score, v = 5)
vini_folds
```



```{r}
lm_spec <-
  linear_reg() %>% 
  set_engine("lm")


ranger_spec <-
  rand_forest(trees = 1e3) %>%
  set_engine("ranger") %>%
  set_mode("regression")
```



```{r}
vini_formula <- score ~ .

lm_wf    <- workflow(vini_formula, lm_spec)
ranger_wf <- workflow(vini_formula, ranger_spec)
```


## Evaluate models


```{r}
contrl_preds <- control_resamples(save_pred = TRUE)

lm_rs <- fit_resamples(
  lm_wf,
  resamples = vini_folds,
  control = contrl_preds
)

ranger_rs <- fit_resamples(
  ranger_wf,
  resamples = vini_folds,
  control = contrl_preds
)
```

How did these two models compare?

```{r}
collect_metrics(lm_rs)
collect_metrics(ranger_rs)
```

The random forest model performs slightly better...

```{r}
final_fitted <- last_fit(ranger_wf, vini_split)
collect_metrics(final_fitted)  ## metrics evaluated on the *testing* data
```

This object contains a fitted workflow that we can use for prediction.

```{r}
final_wf <- extract_workflow(final_fitted)
print(predict(final_wf, vini_test[1,]))
print(vini_test$score[1])
```




